---
title: "Exploratory Analysis Example"
output: html_notebook
---

This is an example of an exploratory analysis being conducted.

source of data: https://archive.ics.uci.edu/ml/datasets/human+activity+recognition+using+smartphones

Obs: The data was cleaned and adjusted with python first.

### Loading Data

```{r}
data <- read.csv('samsungData.csv', sep = ",")
data <- data[ , !(names(data) %in% 'X')]
dim(data)

```

# Plotting average acceleration for first subject

```{r}
#Organizing how the plots will be displayed
par(mfrow= c( 1, 2), mar = c( 5, 4, 1, 1))
#Getting the data and transforming the colum ActivityName into factor
samsungData <- transform(data, ActivityName = factor(ActivityName))
#Subsetting for the subject only
sub1 <- subset(samsungData, subject == 1)
plot(sub1[, 1], col = sub1$ActivityName, ylab = names(sub1)[1])
plot(sub1[, 2], col = sub1$ActivityName, ylab = names(sub1)[2])
legend("bottomright", legend = unique(sub1$ActivityName), col = unique(sub1$ActivityName), pch = 1)

```

### What we can see
The mean body acceleration is uninsteresting for activities as standing, sitting, laying. But it is much more interesting for the others where this measures have more variability.


# Clustering based just on Average Acceleration

```{r}
source("myplclust.R")
distanceMatrix <- dist(sub1[,1:3])
hclustering  <- hclust(distanceMatrix)
myplclust(hclustering, lab.col = unclass(sub1$ActivityName))
```

### What we can see
There is no interesting pattern here to help us to differentiate between the activies

# Plotting max acceleration for the firs subject

```{r}
par(mfrow = c(1,2))
plot(sub1[, 10], pch = 19, col = sub1$ActivityName, ylab = names(sub1)[10])
plot(sub1[, 11], pch = 19, col = sub1$ActivityName, ylab = names(sub1)[11])
```

### What we can see
Again, there is no interesting pattern for passive activities, for activities as standing, sitting, laying that is not the same for the other activies, which present a lot of variability

# Clustering based on maximum acceleration

```{r}
source("myplclust.R")
distanceMatrix <- dist(sub1[, 10:12])
hclustering <- hclust(distanceMatrix)
myplclust(hclustering, lab.col = unclass(sub1$ActivityName))

```
### What we can see
There is a clear pattern for the walking activities in the left hand side, and the non moving in the right side. But whithin these groups, is hard to differentiate the activies yet.

# Singular Value Decomposition

```{r}
svd1 = svd(scale(sub1[, -c(562, 563, 564)]))
par(mfrow = c(1, 2))
plot(svd1$u[, 1], col = sub1$ActivityName, pch = 19)
plot(svd1$u[, 2], col = sub1$ActivityName, pch = 19)
```
6;54

### What we can see
The first matrix seems to separate well the two kinds of activities. The second is not so clear yet.

# Find maximum contributor

Let's try to find what feature is being responsible for the majority of variation for this matrix 
```{r}
plot(svd1$v[,2], pch = 19)
```

# New Clustering with Maximum Contributer

```{r}

maxContrib <- which.max(svd1$v[, 2])
distanceMatrix <- dist(sub1[, c(10:12, maxContrib)])
hclustering <- hclust(distanceMatrix)
myplclust(hclustering, lab.col = unclass(sub1$ActivityName))

```
### What we can see
The movements activities are clearly better separated now

```{r}
names(samsungData)[maxContrib] #The major responsible for the variations
```
### K-means clustering (nstart=1, first try)

```{r}
kClust <- kmeans(sub1[, -c(562,563, 564)], centers = 6)
table(kClust$cluster, sub1$ActivityName)
```
### What we can see
Some of the variables are being well clustered, but we are having problem with some of them yet.

# K-means clustering (nstart = 100, frist try)

```{r}
kClust <- kmeans(sub1[, -c(562,563, 564)], centers = 6, nstart = 100)
table(kClust$cluster, sub1$ActivityName)
```

### What we can see
Walking down, walking, walking up are still better divided, while with the others we did not get much better results'

# Cluster 1 Variable Centers (Laying)

```{r}
plot(kClust$centers[1, 1:10], pch = 19, ylab = "Cluster Center", xlab = "")
```
# Cluster 2 Variable Centers (Laying)

```{r}
plot(kClust$centers[4, 1:10], pch = 19, ylab = "Cluster Center", xlab = "")
```
### What can we see
Based on that, we can see what features are more important for su to find a medium location for our results