---
title: "R Notebook"
output: html_notebook
---

# Asking the question

What are you looking for? What is the hypothesis or question that you are trying to answer with this dataset?
For this data set for example, are air pollution levels lower now then they were before? 

# Getting the data for 1999

The dataset was downloaded from the github repository of the course, and it is in txt format

Let's read the dataframe and show the first lines
```{r, paged.print=FALSE}
pm0 <- read.table(file = "RD_501_88101_1999-0.txt", header = FALSE, sep = "|", na.strings = "", comment.char = "#")

head(pm0)
```
Getting the name of the columns

```{r, , paged.print=FALSE}
#Reading the first line of the file
cnames <- readLines("RD_501_88101_1999-0.txt", 1)
#Spliting the names by the given pattern that is fixed repeated
cnames <- strsplit(cnames, "|", fixed = TRUE)
#Assigning it for the dataframe columns
#Making the names well fitted for column names, since the last ones had spaces in between
names(pm0) <- make.names(cnames[[1]])

```


```{r}
x0 <- pm0$Sample.Value
suma<-summary(x0)
```

```{r}
#Calculating the proportion of missing values
mean(is.na(x0))
```
### Getting data for 2012

```{r, paged.print=FALSE}
pm1 <- read.table(file = "RD_501_88101_2012-0.txt", header = FALSE, sep = "|", na.strings = "", comment.char = "#")

head(pm1)
```
```{r}
#Seeing the dimensions
print(c("Dimensions", dim(pm1)))

names(pm1) <- make.names(cnames[[1]])
x1 <- pm1$Sample.Value
```

### First comparison between the data

```{r}
summary(x0)
summary(x1)
```
### Starting with the visual Representation

```{r}
boxplot(x0, x1)
```
We can see that both that are right skewed. But let's improve the scale for seeing this boxplots
```{r}
boxplot(log10(x0), log10(x1))

```

### Issues with negative numbers

Note how important is to know how the data was collected and what were the procedures. To calculate the amount of dust or pollution in the air, they use a filter, and then they weight the dust that the filter collected. Knowing that, there was not supposed to exist negative values. Let's see what is happening

```{r}
negative <- x1 < 0
sum(negative, na.rm = TRUE)
mean(negative, na.rm = TRUE)
```

### Converting the dates into a date object

Let's start by converting the dates into a object date.

```{r}
#Converting the dates into object dates
dates <- pm1$Date
dates <- as.Date(as.character(dates), "%Y%m%d")

#Plotting the histogram, and showing the format to display the x axisx
hist(dates, "month", format = "%b")

hist(dates[negative], "month", format = "%b")
```

### Analysing a spefic location

```{r}
#Subsetting the data that we want
site0 <- unique(subset(pm0, State.Code == 36, c(County.Code, Site.ID)))
site1 <- unique(subset(pm1, State.Code == 36, c(County.Code, Site.ID)))

#Pasting the information together
site0 <- paste(site0[, 1],site0[, 2], sep = ".")
site1 <- paste(site1[, 1],site1[, 2], sep = ".")

#Look that they are different, so to compare both, we need the intersection between the two vectors of places
str(site1)
str(site0)
both <- intersect(site0, site1)
```
Getting only the data from New York

```{r}
pm0$county.site <- with(pm0, paste(County.Code, Site.ID, sep = "."))
pm1$county.site <- with(pm1, paste(County.Code, Site.ID, sep = "."))
cnt0 <- subset(pm0, State.Code == 36 & county.site %in% both)
cnt1 <- subset(pm1, State.Code == 36 & county.site %in% both)
```

```{r}
#Number of observations in each of the dataframes
sapply(split(cnt0, cnt0$county.site), nrow)
sapply(split(cnt1, cnt1$county.site), nrow)
```

```{r}
pm1sub <- subset(pm1, State.Code == 36 & County.Code == 63 & Site.ID == 2008)
pm0sub <- subset(pm0, State.Code == 36 & County.Code == 63 & Site.ID == 2008)

```

### Now, let's plot the series as a time series

```{r}
dates1 <- as.Date(as.character(pm1sub$Date), format = "%Y%m%d")
x1sub <- pm1sub$Sample.Value
plot(dates1, x1sub)

dates0 <- as.Date(as.character(pm0sub$Date), format = "%Y%m%d")
x0sub <- pm0sub$Sample.Value
plot(dates0, x0sub)
```
### Plotting them together
```{r}
par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
limits <- range(x0sub, x1sub, na.rm = T)
plot(dates0, x0sub, pch = 20, ylim= limits)
abline(h = median(x0sub, na.rm = TRUE))
plot(dates1, x1sub, pch = 20, ylim= limits)
abline(h = median(x1sub, na.rm = TRUE))
```
### Calculating the mean for State
```{r}
mn0 <- with(pm0, tapply(Sample.Value, State.Code, mean, na.rm = T))
summary(mn0)
mn1 <- with(pm1, tapply(Sample.Value, State.Code, mean, na.rm = T))
summary(mn1)
```
```{r echo=TRUE}
d0 <- data.frame(state = names(mn0), mean = mn0)
d1 <- data.frame(state = names(mn1), mean = mn1)
merg <- merge(d0, d1, by = "state")

par(mfrow = c(1, 1))
with(merg, plot(rep(1999, 52), merg[, 2], xlim = c(1998, 2013)))
with(merg, points(rep(2012, 52), merg[, 3]))
segments(rep(1999, 52), merg[,2], rep(2012, 52), merg[, 3])
```


